---
title: "Course Project 1"
author: "Jin Yong Kim"
date: "25 October 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

###Loading and preprocessing the data
This part of the code is necessary for the inital loading of the dataset
```{r}
setwd("c:/Test/repdata")
echo = TRUE
activity <- NULL
activity <- read.csv("activity.csv", header = T, sep = ",")
```

The variable created during by this code are set to NULL.

```{r}
echo = TRUE
df_summary <- NULL
su2 <- NULL
su <- NULL
mn_int <- NULL
activity2 <- NULL
mean_su2 <- NULL
median_su2 <- NULL
activity2_weekend <- NULL
activity2_weekday <- NULL
mean_activity2_weekday <- NULL
mean_activity2_weekend <- NULL
```

### What is mean total number of steps taken per day?
First, the total (sum) of steps is determined for every single date.
```{r}
echo = TRUE
su <- tapply(activity$steps, activity$date, sum, na.rm=T)
```
```{r}
echo = TRUE
hist(su, xlab = "sum of steps per day", main = "histogram of steps per day")
```

The mean and the median total number of steps taken per day are reported :
```{r}
echo = TRUE
mean_su <- round(mean(su))
median_su <- round(median(su))

print(c("The mean is",mean_su))
```

```{r}
print(c("The median is",median_su))
```

### what is the average daily activity pattern ?
A time series plot of the 5-minute interval and the average number of steps taken (averaged across all days) is shown below:
```{r}
echo = TRUE
mn_int <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(mn_int ~ unique(activity$interval), type="l", xlab = "5-min interval")
```

### Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

First, in order to visualize in which variable the NAs are:
```{r}
echo = TRUE
table(is.na(activity) == TRUE)
```
```{r}
summary(activity)
```
There are 2304 missing values.

I will use MICE package to fill out missing values.
```{r}
library(mice)
library(randomForest)
miceMod <- mice(activity, method="rf")
activity2 <- complete(miceMod)
#Check whether there is missing values
colnames(activity2)[colSums(is.na(activity2)) > 0]
```
For comparison with NA's and without (see earlier):
```{r}
echo = TRUE
summary(activity2)
```

### Are there differences in activity patterns between weekdays and weekends?
A new column is added to the dataframe, this column will contain the factor "weekday days"" or "weekend days".
```{r}
echo = TRUE
activity2$weekday <- c("weekday")
activity2[weekdays(as.Date(activity2[, 2])) %in% c("Saturday", "Sunday", "samedi", "dimanche", "saturday", "sunday", "Samedi", "Dimanche"), ][4] <- c("weekend")
table(activity2$weekday == "weekend")
```

```{r}
activity2$weekday <- factor(activity2$weekday)
```
In order to visualize the difference bewteen weekends and days of the week, a new dataframe is created to be usable by the lattice package. First, the data are calculated:
```{r}
echo = TRUE
activity2_weekend <- subset(activity2, activity2$weekday == "weekend")
activity2_weekday <- subset(activity2, activity2$weekday == "weekday")

mean_activity2_weekday <- tapply(activity2_weekday$steps, activity2_weekday$interval, mean)
mean_activity2_weekend <- tapply(activity2_weekend$steps, activity2_weekend$interval, mean)
```
Let's plot it
```{r}
echo = TRUE
library(lattice)
df_weekday <- NULL
df_weekend <- NULL
df_final <- NULL
df_weekday <- data.frame(interval = unique(activity2_weekday$interval), avg = as.numeric(mean_activity2_weekday), day = rep("weekday", length(mean_activity2_weekday)))
df_weekend <- data.frame(interval = unique(activity2_weekend$interval), avg = as.numeric(mean_activity2_weekend), day = rep("weekend", length(mean_activity2_weekend)))
df_final <- rbind(df_weekday, df_weekend)

xyplot(avg ~ interval | day, data = df_final, layout = c(1, 2), 
       type = "l", ylab = "Number of steps")
```


It can be observed that there is a small difference between the period.
       

